---
title: "Detecting Cranial Variation in Modern Populations"
author: "Daniel Cusimano"
date: "04/05/2023"
output: html_document
---
This project is centered around original research from Dr. Kat Platchard (formerly Kat Harrington) conducted for her PhD dissertaion at UH Manoa, 2020. I will use this .html to act as my notes page in the process of cleaning and analyzing the given data. I am not currently sure it this will be able to pull code chunks from two separate codes, but I will find out.

# Goal 1

My first goal for this project will be to consolidate the three spreadsheets provided into a single dataset. An additional need here will be to define all of the measurements in the datasets that are currently marked by acronyms. 

```{r }
#| cache = FALSE
knitr::read_chunk('processingcode.r')
```

```{r, loadpackages}
#| message: FALSE
```

```{r, loaddata}
```

# Data Management

 The above code was used for the previous project. The only issue here is that I have multiple .csv files. When I read data in using the following code, I only pull information from one file (Raw_Japan).


```{r, viewdata1}
```

# Data Dictionary

This first data dictionary contains the definitions of cranial landmarks used in this study.

```{r, datadictionary1}
```
