---
title: "Detecting Cranial Variation in Modern Populations"
subtitle: "Data Processing"
author: "Daniel Cusimano"
date: "04/05/2023"
output: html_document
---
This project is centered around original research from Dr. Kat Platchard (formerly Kat Harrington) conducted for her PhD dissertaion at UH Manoa, 2020. I will use this .html to act as my notes page in the process of cleaning and analyzing the given data. 

This markdown file is mostly for me. I will walk through my logic for each step. 

# Data Processing Layout

My first objective for this project is to prepare the data for analysis. While Dr. Platchard was able to complete various statistical analyses, she did not do so by using the r statistical software. Thus, the rawdata for this project will need to be manipulated prior to its analysis. 

The ultimate goal is to organize the data into an array. Before this, I will organize the files into a simplified format for visualization. 

# Set up for R 

This markdown utilizes codechunks. 

```{r }
#| cache = FALSE
knitr::read_chunk('processingcode.r')
```

A variety of packages are needed. 

```{r, loadpackages}
#| message: FALSE
#| warning: FALSE 
```


```{r, loaddata}
```

# Data Management

I have two primary data files containing 3D craniometric landmark data for populations in East Asia. Before altering any files I need to get a sense of what I'm working with. 

```{r, viewdata1}
```

```{r, viewdata2}
```

```{r, datadictionary1}
```

```{r, datadictionary2}
```

```{r, mergedata}
```

```{r, arrangedata}
```

```{r, d1Array}
```

```{r, d2Array}
```